Climate Prediction Markets and the Convergence of Forecasting Models: An Agent-Based Simulation

### 2016-03-01 paper.Rmd
Work on paper.

---
title: "Climate Prediction Markets and the Convergence of Forecasting Models"
author: " "
date: "Feb 23, 2016"
output: 
  pdf_document:
    keep_tex: true
bibliography: library.bib
---

@@

We present an agent-based model of a climate prediction market. Trader agents adapt their beliefs about the climate based on the monetary performance of their neighboring traders in a social network. We analyze the model to determine how a variety of market and climate factors affect whether climate prediction markets foster a convergence of market participants' beliefs regarding a true model of the global climate.

@@

The previous two decades saw a strong polarization of the climate change debate. Although the scientific consensus on the anthropogenic nature of climate change strongly increased, the average belief about anthropogenic climate change did not evolve much within the public [@Vandenbergh2013f]. In addition, the divide on anthropogenic climate change between liberals and conservatives has grown steadily, indicating that the question is becoming increasingly politicized and potentially disconnected from scientific evidence. This is troubling given the costs of misinformed climate policy. If anthropogenic climate change is a myth but is taken to be true, a tremendous amount of public resources will be wasted on adaptation and mitigation efforts. On the other hand, if climate change is human-induced but not recognized as so, the costs of inaction would be huge. Given that effective climate policies require long-lasting measures to be taken, the importance of an accurate consensus on the issue is manifest.

Attempts to foster such consensus face many social and psychological challenges. Such challenges could be met by setting up climate prediction markets [@Hsu2011; @Vandenbergh2013f]. The idea of using prediction markets to efficiently aggregate information about uncertain events has been discussed for decades [@Horn2014c]. Prediction markets have been shown to have interesting theoretical properties [@Set1998; @Hanson2012] and to perform well in terms of prediction accuracy in experiments [@Hanson2006; @Healy2010], agent-based models (ABM) [@Klingert2012c; @Jumadinova2011a], and the real-world [@Wolfers2006]. We simulate a prediction market for climate outcomes to explore whether, and under what conditions, prediction markets are useful for increasing convergence of climate beliefs.

@@

JG - describe the time series models here.

## Beliefs

Traders use one of the two models to forecast future temperatures in order to trade on the prediction market. These $2$ models are interpreted as the trader's beliefs about the true climate model. They are meant to represent pervasive positions on climate change in the public debate. In order to approximatively match the current configuration of beliefs in climate change in the US, model 1 is randomly assigned to half of the traders, while 2 is assigned to the other half of the traders.

 Thus, both when **true.model** = 0 and when **true.model** = 1,  some traders use the true model to make predictions. However, this does not mean that traders using the true model make perfectly accurate predictions. Although these traders believe in the correct *functional form* of the model, they still need to calibrate it based on limited noisy data. Therefore, the values they use as  the parameters of the model will typically still be different from the true parameters of the data-generating process.

@@

Traders are initially endowed with one single unit Experimental Currency Unit (ECU). Traders use their believed model to forecast the distribution of future temperatures and determine their reservation price for different securities. Securities pay 1 ECU  if the actual temperature at some time $t^*$ falls between a certain range. Traders are assumed to be risk-neutral expected utility maximizers. Therefore, their reservation price for a security $\tilde{s}$ is simply their assessment of the probability that the temperature at time $t^*$ falls in the range covered by $\tilde{s}$.

Based on their reservation price, traders behave as "zero-intelligence" agents [@Gode1993], that is, they sell at a random price above their reservation, and buy at a random price below their reservation. The distribution from which buy and sell prices are drawn is determined by the traders risk attitude, characterized by parameter $risk.tak_i$. These strategies are simple but have been shown to provide good approximations of the behavior on prediction markets (see for instance [@Klingert2012c]), and in financial markets more broadly.

Continuous-double auctions (CDAs) or some variants thereof are common market mechanisms, i.e. procedures to match buy and sell orders. CDA are notably used on large stock markets (see for instance [@Tseng2010a]). Our model features a particular version of CDA through which traders exchange securities.

@@

Traders are part of a social network. Every time a security is realized, each trader $i$ looks at the performance of their neighbors in the network. If one of $i$'s neighbors, say $j$, is richer than $i$, $i$ interprets this has an indication that $j$ has a better approximate model. Then $i$ considers adopting $j$'s approximate model. Traders start with the same initial amount of money, so differences in money among traders can only come from  traders' interactions through the market.} For each trader, the willingness to revise her belief is determined by how ideologically loaded her belief is, which is characterized by parameter $ideo_i$.

 An example of such social network is depicted in Figure X. We assume that the initial network is segregated, that is traders who believe in model 2 (1) are more likely to be linked with other traders who believe in 2 (1). The degree of segregation is determined by parameter **seg**. Although traders can change their approximate model as time passes, the connections between traders do not change as the market unfolds (i.e. the edges are fixed).

@@

The time periods $t$ are grouped in trading *sequences*. In a given sequence, the potential payments associated with traded securities are all based on the temperature at the end of the sequence. For instance, the third trading sequence might start in period $t=1964$ and end in period $t = 1970$. In this case, a security traded in the third sequence pay 1 ECU if the temperature at $t=1970$ falls into the range of temperatures covered by the security.

At each time $t$, traders are assumed to know the past value of the temperature $\hat{T}_{0:t}$ and greenhouse gas emissions  $GHG_{0:t}$. In a sequence finishing at time $t^*$,  traders also have common knowledge of $GHG_{t:t^*}$, the future values of greenhouse gas emissions up to $t^*$. However, at any $t$, traders do not know the value of any future *temperatures*. In particular, in a given sequence, traders do not know the value of $\hat{T}_{t^*}$. Traders can only predict $\hat{T}_{t^*}$ using their approximate model and their knowledge of $\hat{T}_{0:t}$ and  $GHG_{0:t^*}$. Notice that because $GHG_{0:t^*}$ is common knowledge, in each period $t$ every trader $i$ with the same approximate model forms the same pseudo-expectation

 	  $\mu_{ti} \equiv E_t( \hat{T}_{t^*} ~|~ \hat{T}_{0:t}, GHG_{0:t^*}$

 	  $i$'s approximate model at t.

 	We assume that the believed probability distribution of $\hat{T}_{t^*}$ for trader $i$  at any time $t$ is

 	   $\mu_{ti} + \eta_t$

where the distribution $\eta_t$ is obtained by trader $i$ by bootstrapping the residuals from the regression of her believed true model over the past periods. This explains the use of the term pseudo-expectation. Because $\eta_t$ is not bound to have expected value zero, a traders' effective expected value of $\hat{T}_{t^*}$ may be different from $\mu_{ti}$.

@@

The model depends on the following parameters.

@@

The sensitivity analysis is based on a Latin Hypercube Sampling of ... parameter sets from the following distributions [@beachkofski_improved_2002,@carnell_lhs_2012].

@@

We use the model to simulate 10,000 outcomes based on the input paramter sets and then conduct a partial rank correlation coefficient analysis on the relationship between the input matrix, $X$, resulting simulated vector, $y$ [@Marino2008,@pujol_sensitivity:_2014,@saltelli_sensitivity_2009]. Partial correlation computes the linear relationship between the part of the variation of $X_i$ and $y$ that are linearly independent of other $X_j$  ($j \neq i$).  The only difference with the partial correlation and partial rank correlation (which we use here), is that the outcome variable $y$ is first ranked-transformed in order to allow us capture potentially non-linear relationships. That is every value, $y_s$, in the data set is replace by a number $f(y_s) \equiv |\{y_r~|~ y_s > y_r\}|$, where $y_r$ are the other observed values of $y$ in the data set.

Our sensitivity analysis, illustrated by ...,

>>>>

---
title: "Climate Prediction Markets and the Convergence of Forecasting Models: An Agent-Based Simulation"
author: " "
date: "Feb 29, 2016"
output: 
  pdf_document:
    keep_tex: true
bibliography: library.bib
---

@@

We present an agent-based simulation model of a climate prediction market. Traders adapt their beliefs about the future global temperatures based on the monetary performance of their neighboring traders in a social network. We conduct simulation experiments with the model to determine how a variety of market factors and the climate data-generating process affect whether climate prediction markets foster a convergence of market participants' beliefs regarding a true model of the global climate.

@@

The climate change debate has become strongly polarized over the past twenty years. Although the scientific consensus on the anthropogenic nature of climate change strongly increased, the average belief about anthropogenic climate change did not evolve much within the public [@Vandenbergh2013f]. In addition, the divide on anthropogenic climate change between liberals and conservatives has grown steadily as the question is becoming increasingly politicized and potentially disconnected from scientific evidence. The costs of misinformed climate policies are high. If anthropogenic climate change is a myth but is taken to be true, public resources will be spent on unnecessary efforts. On the other hand, if climate change is human-induced but not recognized as so, the costs of inaction could be devastating. Given that effective climate policies require long-lasting measures to be taken, the importance of an accurate consensus on the issue is manifest.

Attempts to foster such consensus face many social and psychological challenges, some of which could be adressed by creating climate prediction markets [@Hsu2011; @Vandenbergh2013f]. The idea of using prediction markets to efficiently aggregate information about uncertain events has been widely discussed [@Horn2014c]. Prediction markets have been shown to have interesting theoretical properties [@Set1998; @Hanson2012] and to perform well in terms of prediction accuracy and information aggregation in experiments [@Hanson2006; @Healy2010], agent-based models (ABM) [@Klingert2012c; @Jumadinova2011a], and the real-world [@Wolfers2006]. We simulate a prediction market where traders exchange securities related to climate outcomes to explore whether, and under what market and climate conditions, prediction markets may be useful for increasing convergence of climate beliefs.

@@

JG - describe the temperature time series models here.

## Climate Beliefs

Traders use one of the two models to forecast future temperatures. These models are interpreted as the trader's beliefs about the true climate process, i.e. the driving factors of long-term global temperatures. They represent pervasive positions on climate change in the public debate. In order to approximatively match the current configuration of beliefs in climate change in the United States, model 1 is randomly assigned to half of the traders, while 2 is assigned to the other half of the traders.

Thus, both when **true.model** = 0 and **true.model** = 1,  some traders use the true data-generating model to make predictions. However, this does not mean that traders using the true model make perfectly accurate predictions. Although these traders believe in the correct *functional form* of the model, they still need to calibrate it based on limited noisy data. Therefore, the values they use as the parameters of the model will typically still be different from the exact parameters of the data-generating process.

@@

Traders are initially endowed with a single unit Experimental Currency Unit (ECU). Traders use their model to forecast the distribution of future temperatures and determine their reservation price for different securities. Securities pay 1 ECU  if the actual temperature at some time $t^*$ falls between a certain range. Traders are risk-neutral expected utility maximizers. Therefore, their reservation price for a security $\tilde{s}$ is simply their assessment of the probability that the temperature at time $t^*$ falls in the range covered by $\tilde{s}$.

Based on their reservation price, traders behave as "zero-intelligence" agents [@Gode1993], that is, they sell at a random price above their reservation, and buy at a random price below their reservation. The distribution from which buy and sell prices are drawn is determined by the traders risk attitude, characterized by parameter $risk.tak_i$. These strategies are simple but have been shown to provide accurate approximations of behavior in prediction markets (see for instance @Klingert2012c), and in financial markets more broadly.

Continuous-double auctions (CDAs) or some variants thereof are common market mechanisms, i.e. procedures to match buy and sell orders. CDA are notably used on large stock markets [@Tseng2010a]. Our model features a particular version of CDA through which traders exchange securities.

@@

Traders are part of a social network. Jonathan - add in changes made to social network. Every time a security is realized, each trader $i$ looks at the performance of their neighbors in the network. If one of $i$'s neighbors, say $j$, is richer than $i$, $i$ interprets this has an indication that $j$ has a better approximate model. Then $i$ considers adopting $j$'s approximate model. Traders start with the same initial amount of money, so differences in money among traders can only come from  traders' interactions through the market. For each trader, the willingness to revise her belief is determined by how ideologically loaded her belief is, which is characterized by parameter $ideo_i$.

An example of such social network is depicted in Figure X. We assume that the initial network is segregated, that is, traders who believe in model 2 (1) are more likely to be linked with other traders who believe in 2 (1). The degree of segregation is determined by parameter **seg**. Although traders can change their approximate model over time, the connections between traders do not change as the market unfolds -- the edges are fixed.

@@

The time periods, $t$, are grouped in trading *sequences*. In a given sequence, the potential payments associated with traded securities are all based on the temperature at the end of the sequence. For instance, the third trading sequence might start in period $t = 1964$ and end in period $t = 1970$. In this case, a security traded in the third sequence pays 1 ECU if the temperature at $t = 1970$ falls into the range of temperatures covered by the security.

At each time $t$, traders are assumed to know the past value of the temperature $\hat{T}_{0:t}$ and greenhouse gas emissions  $GHG_{0:t}$. In a sequence finishing at time $t^*$,  traders also have common knowledge of $GHG_{t:t^*}$, the future values of greenhouse gas emissions up to $t^*$. However, at any $t$, traders do not know the value of any future *temperatures*. In particular, in a given sequence, traders do not know the value of $\hat{T}_{t^*}$. Traders can only predict $\hat{T}_{t^*}$ using their approximate model and their knowledge of $\hat{T}_{0:t}$ and  $GHG_{0:t^*}$. Notice that because $GHG_{0:t^*}$ is common knowledge, in each period $t$ every trader $i$ with the same approximate model forms the same pseudo-expectation $\mu_{ti} \equiv E_t( \hat{T}_{t^*} ~|~ \hat{T}_{0:t}, GHG_{0:t^*})$, $i$'s approximate model at $t$.

We assume that the believed probability distribution of $\hat{T}_{t^*}$ for trader $i$  at any time $t$ is $\mu_{ti} + \eta_t$, where the distribution $\eta_t$ is obtained by trader $i$ by bootstrapping the residuals from the regression of her believed true model over the past periods. This explains the use of the term pseudo-expectation. Because $\eta_t$ is not bound to have expected value zero, a traders' effective expected value of $\hat{T}_{t^*}$ may be different from $\mu_{ti}$.

@@

The model depends on the following parameters, which we vary in simulation experiments to determine their effects on the convergence of beliefs.

@@

The sensitivity analysis is based on a Latin Hypercube Sampling of parameter sets from the following distributions [@beachkofski_improved_2002,@carnell_lhs_2012].

@@

We use the model to simulate 10,000 outcomes based on the input paramter sets and then conduct a partial rank correlation coefficient analysis on the relationship between the input matrix, $X$, resulting simulated vector, $y$ [@Marino2008; @pujol_sensitivity:_2014; @saltelli_sensitivity_2009]. Partial correlation computes the linear relationship between the part of the variation of $X_i$ and $y$ that are linearly independent of other $X_j$  ($j \neq i$).  The only difference with the partial correlation and partial rank correlation (which we use here), is that the outcome variable $y$ is first ranked-transformed in order to allow us capture potentially non-linear relationships. That is every value, $y_s$, in the data set is replace by a number $f(y_s) \equiv |\{y_r~|~ y_s > y_r\}|$, where $y_r$ are the other observed values of $y$ in the data set.

Our sensitivity analysis, Fig X.
